Let's talk about Thurstone & Co.: An information-theoretical model for comparative judgments, and its statistical translation

### 2024-11-21 index.qmd
section:issues; process:write

<!-- 2. What is the effectiveness of CJ -->
Numerous studies have documented the effectiveness of CJ in assessing traits and competencies over the past decade. These studies have emphasized three aspects of the method's effectiveness: its reliability, validity, and practical applicability. Research on reliability indicates that CJ requires a relatively small number of pairwise comparisons [@Verhavert_et_al_2019; @Crompvoets_et_al_2022] to produce trait scores that are as precise and consistent as those generated by other assessment methods [@Coertjens_et_al_2017; @Goossens_et_al_2018; @Bouwer_et_al_2023]. Furthermore, evidence suggests that the reliability and time efficiency of CJ are comparable, if not superior, to those of other assessment methods when employing adaptive comparison algorithms [@Pollitt_2012b; @Verhavert_et_al_2022; @Mikhailiuk_et_al_2021]. On the other hand, research on validity suggests that scores generated by CJ can accurately represent the traits under measurement [@Whitehouse_2012; @vanDaal_et_al_2016; @Lesterhuis_2018_thesis; @Bartholomew_et_al_2018; @Bouwer_et_al_2023]. Finally, research on practical applicability highlights the method's versatility across both educational and non-educational contexts [@Kimbell_2012; @Jones_et_al_2015; @Bartholomew_et_al_2018; @Jones_et_al_2019; @Marshall_et_al_2020; @Bartholomew_et_al_2020; @Boonen_et_al_2020].

@@

<!-- 3. But what critical issues remain CJ Research -->
Nevertheless, despite the increasing number of CJ studies, unsystematic and fragmented research approaches have left several critical issues unaddressed. This research primarily focuses on three: the over-reliance on Thurstone's Case V assumptions in the statistical analysis of CJ data, the apparent disconnect between CJ's trait measurement and hypothesis testing, and the unclear role of comparison algorithms on the method's reliability and validity. The following sections will discuss each of these issues in detail, followed by the introduction of a theoretical model and its statistical translation, which aims to address all three concerns simultaneously.

@@

<!-- 1. Thurstone's Theory of CJ: discriminal process -->
In its most general form, Thurstone's theory [-@Thurstone_1927b] posits that the dichotomous outcome resulting from comparing two stimuli is determined by two factors: the discriminal process of each stimulus and their discriminal difference. The *discriminal process* refers to the psychological effect each stimulus has on the judges, or more simply stated, the judges’ perception of the trait level of each stimulus. Thurstone assumes that the discriminal process for each stimulus follows a Normal distribution. In this distribution, the mode (mean), known as the *modal discriminal process*, represents the position of the stimulus on the trait continuum, while the dispersion, known as the *discriminal dispersion*, reflects the variability in the stimulus' perceived trait level. @fig-discriminal_process shows example distributions of discriminal process for two stimuli (objects).

![Example distributions of discriminal processes for two stimuli (objects). Extracted from @Bramley_2008 [pp. 249]. ](/images/figures/discriminal_process.png){#fig-discriminal_process width=70%}

<!-- 2. Thurstone's Theory of CJ: discriminal difference -->
However, since the discriminal mode and dispersion of a single stimulus are not directly observable except through comparison, the *law of comparative judgment* becomes essential. This law asserts that when assessing a specific trait by comparing two stimuli, the stimulus positioned further along the continuum is perceived as having a higher level of that trait. Thus, the observed dichotomous outcome is determined by the distribution of the difference between the stimuli's discriminal processes, called the *discriminal difference*. @fig-comparative_judgment shows an example distribution of the discriminal difference for two stimuli (objects).

![Distribution of the discriminal difference between two stimuli (objects). Extracted from @Bramley_2008 [pp. 251]. ](/images/figures/comparative_judgment.png){#fig-comparative_judgment width=65%}

<!-- 3. Thurstone's Cases and their simplifying assumptions  -->
Importantly, the theory's general form primarily addresses pairwise comparisons of stimuli made by a single judge [@Thurstone_1927b, pp. 267]. Consequently, to enhance its practical applicability, Thurstone introduced five distinct cases, each defined by progressively simplifying assumptions. @tbl-thurstone_cases summarizes these cases, focusing on key assumptions such as the distribution of discriminal processes, the similarity of discriminal dispersions across stimuli, the correlation between stimuli, and which judges perform the comparisons. For a comprehensive discussion of this progression, refer to @Thurstone_1927b and @Bramley_2008 [pp. 248-253].

## The Case V and the statistical analysis of CJ data

<!-- recording: Sven and Tine 24.10.25; time: 00:31:10 - 00:52:20 -->
<!-- recording: Sven and Tine 24.11.12; time: 00:00:00 - 00:00:00 -->

<!-- 1. The Case V and its simpler statistical representation: the BTL model -->
Surprisingly, despite its reliance on the largest number of simplifying assumptions [@Bramley_2008, pp. 253; @Kelly_et_al_2022, pp. 677], Case V remains the most widely used case in the CJ literature. This popularity is largely due to its simplified statistical representation in the Bradley-Terry-Luce (BTL) model [@Bradley_et_al_1952; @Luce_1959]. The BTL model mirrors Case V’s assumptions, with one key difference: while Case V assumes a Normal distribution for the stimuli's discriminal processes, the BTL model uses the more mathematically tractable Logistic distribution [@Andrich_1978; @Bramley_2008, pp. 254] (see @tbl-thurstone_cases). This substitution has little effect on the model's estimation or interpretation, as the Normal and Logistic distributions differ by a scaling factor of approximately $1.7$ [@vanderLinden_et_al_2017_I, pp. 16] (refer to @fig-logistic_vs_normal).

@@

![Probability density](/images/figures/density.png){#fig-density width=95%}

![Cummulative probability](/images/figures/cumulative.png){#fig-cumulative width=95%}

@@

<!-- 2. But what are limitations of Case V (and the BTL model) -->
However, Case V (and its BTL model counterpart) was originally developed to provide a "rather coarse scaling" of traits [@Thurstone_1927b, pp. 269], prioritizing statistical simplicity over precision in trait measurement [@Kelly_et_al_2022, pp. 677]. As a result, its assumptions may not suit applications beyond the psycho-physical contexts for which it was created. Thurstone himself cautioned that its use "should not be made without (an) experimental test" [@Thurstone_1927b, pp. 270], acknowledging that some assumptions could prove problematic with more complex, less homogeneous stimuli, such as handwriting or English compositions [@Thurstone_1927a, pp. 374]. Consequently, given that current CJ applications often deal with complex, less homogeneous stimuli, two key assumptions of Case V may not consistently hold in theory or practice: the zero correlation and equal dispersion between stimuli.

@@

<!-- 3. What does the zero correlation between stimuli imply? -->
The assumption of *zero correlation between stimuli* is best illustrated with an example. For instance, when evaluating text quality, this assumption suggests that a judge’s perception of quality in one text does not influence their perception of the same trait in the comparison text. Thurstone attributes this independence to the cancellation of potential judges' biases, driven by opposing and equally weighted factors; commonly known as 'mood' and 'simultaneous contrast' effects, that operate during pairwise comparisons [@Thurstone_1927b, pp. 268]. This cancellation of bias has been mathematically demonstrated by @Andrich_1978, assuming additive biases and a logit scale derived from the BTL model.

<!-- 4. But what are the problems of assuming zero correlation? -->
However, given that evidence has shown judges' quality assessments are influenced by multiple, complex aspects of the texts [@vanDaal_et_al_2016; @Lesterhuis_2018; @Chambers_et_al_2022], it is plausible that the accuracy of these judgments also varies based on factors such as age, culture, education, expertise [@Kelly_et_al_2022, pp. 683], and ultimately, based on the individual judge themselves [@Gill_et_al_2013; @vanDaal_et_al_2017; @vanDaal_2020].

[This means that the cancellation of judges' biases may not hold in theory or practice, and that judges' biases may depend on the stimuli, by means of an interaction (but present) or because the judges works as a confounder ]{style="color:red;"}

Such differences may not be detected through analyses of bias and misfit [@Kelly_et_al_2022, pp. 683].

<!-- Would the aggregate view of young, British-Asian men always be the same as the aggregate view of older, black women? [@Kelly_et_al_2022, pp. 683] -->

<!-- The accuracy of judgment may vary by subject and by judge [@Gill_et_al_2013; @vanDaal_et_al_2017; @vanDaal_2020] -->

<!-- What we are saying is that the judges' effects depend on the stimuli, by means of an interaction (but present) or because the judges works as a confounder. This, in the best scenario, cause less precise estimates, and in the worst scenario, biased estimates. possibility of bias is hinted in @Bramley_2008, and later mentioned again in @Kelly_et_al_2022, and some evidence is found in @Pollitt_et_al_2003, @vanDaal_et_al_2017 and @Gijsen_et_al_2021 -->

@@

@@

<!-- this also translates into the idea that stimuli are the main focus of estimation and analysis, but what happens when the focus of analysis is the individuals that generated those stimuli. Meaning there is an amount of correlation that it is not accounted for. Use example of Boonen!! -->

<!--   Here we loose precision, because we are not controlling for clustering -->

<!-- This is the case of clustering or measurement error, and overlooking issues such as clustering and measurement error can lead to biased and less precise parameter estimates [@McElreath_2020], ultimately diminishing the statistical power of models and increasing the likelihood of committing type I or type II errors when addressing research inquiries. -->

@@

## The disconnect between trait measurement and hypothesis testing

@@

<!-- 2. Utilizing BTL Scores for further analysis and hypothesis testing -->
Researchers then typically use the estimated BTL scores, or their transformations, to conduct additional analyses or hypothesis tests. For example, these scores have been used to identify 'misfit' judges and stimuli [@Pollitt_2012b; @vanDaal_et_al_2017; @Goossens_et_al_2018], detect biases in judges' ratings [@Pollitt_et_al_2003; @Pollitt_2012b], calculate correlations with other assessment methods [@Goossens_et_al_2018; @Bouwer_et_al_2023], or test hypotheses related to the underlying trait of interest [@Bramley_et_al_2019; @Boonen_et_al_2020; @Bouwer_et_al_2023; @vanDaal_et_al_2017; @Jones_et_al_2019; @Gijsen_et_al_2021].

<!-- 3. But what are the challenges in using BTL scores for further analysis? -->
However, the statistical literature advises caution when using estimated scores to conduct additional analyses or hypotheses tests. A key consideration is that BTL scores are parameter estimates that inherently carry uncertainty. Ignoring this uncertainty can introduce bias into the analysis and reduce the precision of hypothesis tests. Notably, the direction and magnitude of the bias are often unpredictable; results may be attenuated, exaggerated, or remain unaffected, depending on the amount of uncertainty present in the scores and the actual effects being tested [@Kline_et_al_2023, pp. 25; @Hoyle_et_al_2023, pp. 137]. Furthermore, reduced precision in hypothesis tests weakens their statistical power, ultimately increasing the likelihood of committing type-I or type-II errors [@McElreath_2020]. 
<!-- A type-I error occurs when a *true* null hypothesis is incorrectly rejected, while a type-II error occurs when a *false* null hypothesis is incorrectly accepted [@Everitt_et_al_2010]. -->

@@

## The role and impact of comparison algorithms

@@

# Theory {#sec-theory}

@@

## A theoretical model for CJ {#sec-theory-theoretical}

@@

@@

**Non-financial interests:** Author XX serve on advisory broad of Company Y but receives no compensation this role.

@@

**CRediT authorship contribution statement:** *Conceptualization:* S.G., S.DM., T.vD., and J.M.R.E; *Methodology:* S.DM., T.vD., and J.M.R.E; *Software:* J.M.R.E.; *Validation:* J.M.R.E.; *Formal Analysis:* J.M.R.E.; *Investigation:* J.M.R.E; *Resources:* S.G., S.DM., and T.vD.; *Data curation:* J.M.R.E.; *Writing - original draft:* J.M.R.E.; *Writing - review & editing:* S.G., S.DM., and T.vD.; *Visualization:* J.M.R.E.; *Supervision:* S.G. and S.DM.; *Project administration:* S.G. and S.DM.; *Funding acquisition:* S.G. and S.DM.

>>>>

<!-- 2. What is the effectiveness of CJ -->
Numerous studies have documented the effectiveness of CJ in assessing traits and competencies over the past decade. These studies have emphasized three aspects of the method's effectiveness: its reliability, validity, and practical applicability. Research on reliability indicates that CJ requires a relatively small number of pairwise comparisons [@Verhavert_et_al_2019; @Crompvoets_et_al_2022] to produce trait scores that are as precise and consistent as those generated by other assessment methods [@Coertjens_et_al_2017; @Goossens_et_al_2018; @Bouwer_et_al_2023]. Furthermore, evidence suggests that the reliability and time efficiency of CJ are comparable, if not superior, to those of other assessment methods when employing adaptive comparison algorithms [@Pollitt_2012b; @Verhavert_et_al_2022; @Mikhailiuk_et_al_2021]. Meanwhile, research on validity suggests that scores generated by CJ can accurately represent the traits under measurement [@Whitehouse_2012; @vanDaal_et_al_2016; @Lesterhuis_2018_thesis; @Bartholomew_et_al_2018; @Bouwer_et_al_2023], while research on practical applicability highlights the method's versatility across both educational and non-educational contexts [@Kimbell_2012; @Jones_et_al_2015; @Bartholomew_et_al_2018; @Jones_et_al_2019; @Marshall_et_al_2020; @Bartholomew_et_al_2020; @Boonen_et_al_2020].

@@

<!-- 3. But what critical issues remain in CJ Research?, what is the organization of the study? -->
Nevertheless, despite the increasing number of CJ studies, unsystematic and fragmented research approaches have left several critical issues unaddressed. The present study primarily focuses on three: the over-reliance on the assumptions of Thurstone's Case V in the statistical analysis of CJ data, the apparent disconnect between CJ's trait measurement and hypothesis testing, and the unclear role of the diverse assessment design features on CJ's reliability and validity. The following sections begin with a brief overview of Thurstone's theory and a detailed discussion of these issues. Subsequently, the study introduces a theoretical model for CJ that builds upon Thurstone's theory, alongside its statistical translation, designed to address all three concerns simultaneously.

# Thurstone's theory {#sec-thurstone_theory}
<!-- recording: Sven and Tine 24.11.18; time: 00:04:30 - 00:14:50 -->
<!-- recording: Sven and Tine 24.11.18; time: 00:19:04 - 00:21:37 -->

<!-- 1. Thurstone's Theory of CJ: discriminal process -->
In its most general form, Thurstone's theory [-@Thurstone_1927b] suggests that two factors determine the dichotomous outcome of pairwise comparisons: the discriminal process of each stimulus and their discriminal difference. The *discriminal process* refers to the psychological effect each stimulus exerts on the judges, or more simply, the underlying perception of the stimulus' trait level. According to the theory, the discriminal process for each stimulus follows a Normal distribution. The mode (mean) of this distribution, referred to as the *modal discriminal process*, represents the stimulus' position on the trait continuum. Meanwhile, the dispersion of the distribution, referred to as the *discriminal dispersion*, reflects the variability in the perceived trait level of the stimulus.

<!-- 2. Thurstone's Theory of CJ: discriminal difference -->
However, since the discriminal process of a single stimulus is not directly observable, the *law of comparative judgment* becomes essential. This law states that in pairwise comparisons, the stimulus positioned further along the trait continuum is perceived as having a higher level of that trait. Thus, the theory assumes the observed dichotomous outcome is determined by the distribution of the difference between the underlying discriminal processes of the stimuli, referred to as the *discriminal difference*. This suggests that the outcome depends on the relative distance between stimuli, rather than their absolute positions on the trait continuum.

<!-- 3. Example-->
These concepts can be illustrated through an example. For instance, in the context of evaluating text quality, @fig-discriminal_process could depict the underlying discriminal process distributions for two written texts, highlighting differences in their discriminal dispersions and modal discriminal processes along the quality trait continuum. Furthermore, @fig-discriminal_difference could display the discriminal difference distribution for these texts, showing that text A is perceived to exhibit significantly higher quality than text B, as indicated by the shaded gray area. Consequently, the dichotomous outcome of this comparison would favor text A.

::: {#fig-thurstone_theory layout-ncol=2}

![Discriminal processes](/images/figures/discriminal_process.png){#fig-discriminal_process width=95%}

![Discriminal difference](/images/figures/comparative_judgment.png){#fig-discriminal_difference width=100%}

Example distribution of discriminal processes and their discriminal difference for two written texts (stimuli or objects). Extracted from @Bramley_2008 [pp. 249-251].
:::

<!-- 4. Thurstone's Cases and their simplifying assumptions  -->
Importantly, the general form of Thurstone's theory primarily addressed pairwise comparisons of stimuli made by a single judge [@Thurstone_1927b, pp. 267]. Thus, for practical application, Thurstone introduced five distinct cases derived from this general form, each defined by progressively simplifying assumptions. @tbl-thurstone_cases summarizes these cases, highlighting key assumptions such as the distribution of discriminal processes, the similarity of discriminal dispersions across stimuli, the correlation between stimuli, and the number of judges performing the comparisons. For a comprehensive discussion of this progression, refer to @Thurstone_1927b and @Bramley_2008 [pp. 248-253].

@@

## The Case V and the statistical analysis of CJ data {#sec-theory-issue1}

<!-- recording: Sven and Tine 24.10.25; time: 00:31:10 - 00:52:20 -->
<!-- recording: Sven and Tine 24.11.12; time: 00:00:00 - 00:00:00 -->
<!-- recording: Sven and Tine 24.11.18; time: 00:00:25 - 00:04:30 -->
<!-- recording: Sven and Tine 24.11.18; time: 00:14:50 - 00:19:04 -->
<!-- recording: Sven and Tine 24.11.18; time: 00:21:37 - 00:47:20 -->

<!-- 1. The Case V and its simpler statistical representation: the BTL model -->
Despite its reliance on the largest number of simplifying assumptions [@Bramley_2008, pp. 253; @Kelly_et_al_2022, pp. 677], Case V remains the most widely used case in the CJ literature. This popularity is largely due to its simplified statistical representation in the Bradley-Terry-Luce (BTL) model [@Bradley_et_al_1952; @Luce_1959]. The BTL model mirrors the assumptions of Case V, with one key difference: while Case V assumes a Normal distribution for the discriminal processes of stimuli, the BTL model uses the more mathematically tractable Logistic distribution [@Andrich_1978; @Bramley_2008, pp. 254] (see @tbl-thurstone_cases). This substitution has little impact on the model's estimation or interpretation, as the Normal and Logistic distributions differ by a scaling factor of approximately $1.7$ [@vanderLinden_et_al_2017_I, pp. 16] (see @fig-logistic_vs_normal).

@@

![Probability density](/images/figures/density.png){#fig-density width=85%}

![Cummulative probability](/images/figures/cumulative.png){#fig-cumulative width=85%}

@@

<!-- 2. But what are limitations of Case V (and the BTL model)? -->
However, Case V was originally developed to provide a "rather coarse scaling" of traits [@Thurstone_1927b, pp. 269], prioritizing statistical simplicity over precision in trait measurement [@Kelly_et_al_2022, pp. 677]. As a result, its assumptions may not be suitable for applications beyond the psycho-physical contexts for which it was created. Thurstone himself cautioned that its use "should not be made without (an) experimental test" [@Thurstone_1927b, pp. 270], acknowledging that some assumptions could prove problematic in the presence of complex traits or heterogeneous stimuli, such as handwriting or English compositions [@Thurstone_1927a, pp. 374]. Consequently, given that modern CJ applications frequently involve these types of traits and stimuli, two main assumptions of Case V may not consistently hold in theory or practice: the zero correlation and equal dispersion between stimuli.

@@

<!-- 3. What does the zero correlation between stimuli imply? and does it hold? -->
The assumption of *zero correlation between stimuli* is more easily understood through an example. For instance, when using pairwise comparisons to evaluate text quality, the assumption suggests that a judge's perception of the trait in one text does not influence his perception of the same trait on another. Thurstone attributed this independence to the cancellation of potential judges' biases, driven by two opposing and equally weighted effects occurring during the pairwise comparisons [@Thurstone_1927b, pp. 268]. This cancellation was mathematically demonstrated by @Andrich_1978, assuming discriminal processes with additive biases. However, it is easy to imagine two scenarios where the zero correlation assumption almost certainly does not hold: when the pairwise comparison involves multidimensional, complex traits with heterogeneous stimuli, and when an additional hierarchical structure is relevant to the stimuli.

<!-- 4. The first scenario where the assumption does not hold -->
In the first scenario, evidence suggests that when judges are confronted with multi-dimensional, complex traits and heterogeneous stimuli, such as when evaluating written text quality, their assessments are influenced by multiple, intricate aspects of the stimuli [@vanDaal_et_al_2016; @Lesterhuis_2018; @Chambers_et_al_2022]. In such cases, it is not inconceivable that these various aspects of the stimuli unevenly influence the judges’ perception of one text over another, leading to biases that resist cancellation, as the effects of the factors involved in the comparison are neither equally weighted nor opposing. For example, this may occur when a judge, while assessing the argumentative quality of a text, places disproportionate emphasis on grammatical errors, ultimately favoring texts with fewer errors but weaker arguments. Although the literature does not provide direct evidence for this specific argument, studies such as @Pollitt_et_al_2003 do offer evidence of judges' biases, supporting the idea that the factors involved in pairwise comparisons may not always cancel out.

<!-- research shows that, in the presence of these traits and stimuli, the accuracy of the judges' trait perception is influenced by the rank-order distance of the stimuli [@vanDaal_et_al_2017; @Gijsen_et_al_2021], a coarse measure of the relative positioning of the stimuli within the trait continuum. -->

<!-- 5. The second scenario where the assumption does not hold -->
<!-- In the second scenario, the literature has not even considered the issues related with the existence of an additional hierarchical structure relevant to the stimuli, such as in cases where data includes multiple samples of stimuli within different individuals. In this scenario, there is an amount of correlation that it is not accounted for -->

<!-- 6. But what are the problems when the zero correlation assumption does not hold? -->
<!-- The best the literature has done is to estimate scores and then use those scores to run a multilevel model as in @Boonen_et_al_2020, which in itself brings its own problems. Refer to @sec-theory-issue2 -->

<!-- Such differences may not be detected through analyses of bias and misfit [@Kelly_et_al_2022, pp. 683].  -->

@@

<!-- Cancellation of judges: What we are saying is that the judges' effects depend on the stimuli, by means of an interaction (but present) or because the judges works as a confounder. This, in the best scenario, cause less precise estimates by ignoring clusters, and in the worst scenario, biased estimates if the judges do register biases. -->

<!-- Multilevel structure: Here we loose precision, because we are not controlling for clustering, and overlooking issues such as clustering and measurement error can lead to biased and less precise parameter estimates [@McElreath_2020], ultimately diminishing the statistical power of models and increasing the likelihood of committing type I or type II errors when addressing research inquiries. -->

<!-- (This a is a validity contention) -->
<!-- Furthermore, it is not inconceivable that the selection of a particular group of judges used to perform the pairwise comparison could result in an incomplete depiction of the dimensions and complexity of the trait. Previous research has highlighted that factors such as age, culture, and education [@Kelly_et_al_2022, pp. 683], as well as individual differences among judges [@Gill_et_al_2013; @vanDaal_et_al_2017; @vanDaal_2020], could influence judgment accuracy, thereby affecting the "shared consensus" of the trait measurement. As @Kelly_et_al_2022 noted [pp. 683], “Would the aggregate view of young, British-Asian men always be the same as the aggregate view of older, black women?” -->

@@

@@

## The disconnect between trait measurement and hypothesis testing {#sec-theory-issue2}

@@

<!-- 2. Utilizing BTL Scores for further analysis and hypothesis testing -->
Researchers then typically use the estimated BTL scores, or their transformations, to conduct additional analyses and tests, or to make decisions regarding the exclusion of certain data in these analyses and tests. The literature shows that these scores have been employed to calculate correlations with other assessment methods [@Goossens_et_al_2018; @Bouwer_et_al_2023] or to test hypotheses related to the underlying traits of interest [@Bramley_et_al_2019; @Boonen_et_al_2020; @Bouwer_et_al_2023; @vanDaal_et_al_2017; @Jones_et_al_2019; @Gijsen_et_al_2021]. Additionally, the BTL scores have been used to detect biases in judges' ratings [@Pollitt_et_al_2003; @Pollitt_2012b], as well as to identify "misfit" judges and stimuli [@Pollitt_2012b; @vanDaal_et_al_2017; @Goossens_et_al_2018], with considerations for their possible exclusion.

<!-- 3. But what are the challenges in using BTL scores for further analysis? -->
However, the statistical literature advises caution when using estimated scores to conduct additional analyses and tests while warning against eliminating data through ad-hoc univariate procedures. A key consideration is that BTL scores are parameter estimates that inherently carry uncertainty. Ignoring this uncertainty can bias the analysis and reduce the precision of hypothesis tests. Notably, the direction and magnitude of such biases are often unpredictable. Results may be attenuated, exaggerated, or remain unaffected depending on the degree of uncertainty in the scores and the actual effects being tested [@Kline_et_al_2023, pp. 25; @Hoyle_et_al_2023, pp. 137]. Moreover, eliminating data using ad-hoc univariate procedures using these scores compounds this issue by excluding possible valuable information from the analyses [@McElreath_2020], further introducing bias. Finally, the reduced precision in hypothesis tests weakens their statistical power, increasing the likelihood of type-I or type-II errors [@McElreath_2020].

<!-- A type-I error occurs when a *true* null hypothesis is incorrectly rejected, while a type-II error occurs when a *false* null hypothesis is incorrectly accepted [@Everitt_et_al_2010]. -->

@@

## Assessment design features and their role on reliability and validity {#sec-theory-issue3}

<!-- recording: Sven and Tine 24.11.18; time: 00:47:20 - 00:56:41 -->

@@

# An updated theoretical and statistical model for CJ {#sec-theory}

@@

## The theoretical model {#sec-theory-theoretical}

@@

<!-- This means that the cancellation of judges' biases may not hold in theory or practice, and that judges' biases may depend on the stimuli, by means of an interaction (but present) or because the judges works as a confounder -->

@@

**Non-financial interests:** The authors have no relevant non-financial interest to disclose.

@@

**CRediT authorship contribution statement:** *Conceptualization:* S.G., S.DM., T.vD., and J.M.R.E; *Methodology:* S.DM., T.vD., and J.M.R.E; *Software:* J.M.R.E.; *Validation:* J.M.R.E.; *Formal Analysis:* J.M.R.E.; *Investigation:* J.M.R.E; *Resources:* S.G., S.DM., and T.vD.; *Data curation:* J.M.R.E.; *Writing - original draft:* J.M.R.E.; *Writing - review and editing:* S.G., S.DM., and T.vD.; *Visualization:* J.M.R.E.; *Supervision:* S.G. and S.DM.; *Project administration:* S.G. and S.DM.; *Funding acquisition:* S.G. and S.DM.
